Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-07-01 sattose/paper.tex
Shortened references a bit

\section{Introduction}
Duplication in source code is often seen as one of the most harmful types of technical debt. In Martin Fowler's ``Refactoring'' book \cite{fowler1999refactoring}, he claims that \textit{``Number one in the stink parade is duplicated code. If you see the same code structure in more than one place, you can be sure that your program will be better if you find a way to unify them.''}.
% In this research, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve the maintainability of the refactored code.

@@

A survey by Roy et al. \cite{roy2007survey} describes various ways in which clones can be identified. Most clone detection tools focus on finding clones that align with these definitions. In this paper, we outline challenges with these clone type definitions when considered in a refactoring context. We next propose solutions to these problems that would enable the detection of clones that should be refactored, rather than fragments of code that are just similar.

@@

Our end goal is to improve upon the current state-of-the-art in clone research \cite{fontana2015duplicated, alwaqfi2017refactoring} by building a clone refactoring tool that automatically applies refactorings to a large percentage of clones found. The design decisions for this tool are made on basis of data gathered from a large corpus of software systems together with our own experience and findings from literature.

Section \ref{chap:background} describes the background from literature that we used to conduct this study. Section \ref{chap:clonedetection} presents our clone refactoring tool proposal. Section \ref{chap:clonetypes} outlines challenges with clone type definitions and proposes solutions for them in a refactoring context. Section \ref{chap:clonecontextexpl} outlines our analysis and measurements regarding the context of code clones. Section \ref{chap:mergingdups} shows our measurements regarding the amount of clones that can be refactored through method extraction. Section \ref{chap:threatstovalidity} shows the threats to validity of this study and section \ref{chap:conclusion} draws a conclusion based on our preliminary results.

@@

\subsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be considered for detection. We focus on clone classes, because of the advantages for refactoring. Clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy \cite{fontana2012duplicated}.

\subsection{Clone types}\label{chap:backgroundclonetypes}
In a 2007 survey by Roy et al. \cite{roy2007survey} he defines several types of clones:\\
\textbf{Type 1:} Identical code fragments except for variations in whitespace (may also be variations in layout) and comments.\\
\textbf{Type 2:} Structurally/syntactically identical fragments except for variations in identifiers, literals, types, layout and comments.\\
\textbf{Type 3:} Copied fragments with further modifications. Statements can be changed, added or removed in addition to variations in identifiers, literals, types, layout, and comments.\\

A higher type of clone means that it is harder to detect. It also makes the clone harder to refactor, as more transformations would be required. Higher clone types also become more disputable whether they actually indicate a harmful anti-pattern (as not every clone is harmful \cite{jarzabek2010clones, kapser2008cloning}).

@@

\subsection{Related work in clone refactoring tools}
The Duplicated Code Refactoring Advisor (DCRA) looks into refactoring opportunities for clone pairs \cite{fontana2012duplicated, fontana2015duplicated}. DCRA only focuses on refactoring clone pairs, with the authors arguing that \textit{clone pairs are much easier to manage when considered singularly.} As intermediate steps, the authors measure a corpus of Java systems for some clone-related properties of the systems, like the relation (in terms of inheritance) between code fragments in a clone pair. We further look into these measurements in section \ref{chap:relationsinstances}.

Aries \cite{higo2004aries, higo2008metric} focuses on the detection of refactorable clones. They do this based on the relation between clone instances through inheritance, similar to Fontana et al. \cite{fontana2012duplicated}. Aries also proposes a refactoring: if two clone classes are siblings of each other (share the same superclass), they propose to perform ``Extract method'' and ``Pull up method'' sequentially. This tool only proposes the refactoring, and does not provide help in the process of applying the refactoring.

We investigated several research efforts that look into code clone refactoring \cite{alwaqfi2017refactoring, chen2018clone, koni2001scenario}. However, all of these tools only support a subset of all harmful clones that are found. Also, these tools are limited to suggesting refactoring opportunities, rather than actually applying refactorings where suitable. Finally, all published approaches have limitations, such as false positives in their clone detection \cite{chen2018clone} or being limited to clone pairs \cite{higo2008metric}.

\section{Addressing problems with clone type definitions}\label{chap:clonetypes}
For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings (see section \ref{chap:backgroundclonetypes} for these definitions) to increase the chance that we can refactor the clone while improving the design.

\section{Shortcomings of clone types}
The clone definitions, as outlined in section \ref{chap:backgroundclonetypes}, allow reasoning about the duplication in a software system. Clones by these definitions can relatively easily and efficiently be detected. This has allowed for large scale analyses of duplication. However, these clone type definitions have shortcomings which make the clones detected in correspondence with these definitions less valuable for (automated) refactoring purposes.

@@

\subsection{Type 1 clones} \label{sec:type1}
Type 1 clones are \textit{identical clone fragments except for variations in whitespace and comments} \cite{roy2007survey}. This allows for the detection of clones that are the result of copying and pasting existing code, along with other reasons why duplicates might get into a codebase.

@@

\subsection{Type 2 clones}\label{sec:type2}
Type 2 clones are \textit{structurally/syntactically identical fragments except for variations in identifiers, literals, types, layout and comments} \cite{roy2007survey}. This definition allows for the reasoning about code fragments that were copied and pasted, and then slightly modified. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones. Merging such clones can be harmful instead of helpful.

@@

\subsection{Type 3 clones}\label{sec:type3}
Type 3 clones are \textit{copied fragments with further modification (having added, removed or changed statements)} \cite{roy2007survey}. Detection of clones by this definition can be hard, as it may be hard to detect whether a fragment was copied in the first place if it was severely changed. Because of this, most clone detection implementations of type 3 clones work on basis of a similarity threshold \cite{roy2008nicad,ragkhitwetsagul2019siamese,jiang2007deckard,semura2017ccfindersw}. This similarity threshold has been implemented in different ways: textual similarity (for instance using levenshtein distance) \cite{lavoie2011automated}, token-level similarity \cite{sajnani2016sourcerercc} of statement-level similarity \cite{kamalpriya2017enhancing}.

@@

\subsubsection{T3R clones}\label{sec:type3r}
Type 3 clones allow any change in statements (added, removed and changed statements). Many clone detection tools have implemented this type by introducing a similarity threshold \cite{roy2008nicad,ragkhitwetsagul2019siamese,jiang2007deckard,semura2017ccfindersw}. This similarity threshold entails that two fragments may differ by a certain percentage. This means that T3R allows the inclusion of a statement that is not detected by type 1 or 2 clone detection. When looking at how we can refactor a statement that is not included by one clone instance but is in another, we find that we require a conditional block to make up for the difference in statements. This is a tradeoff, as an added conditional block increases the complexity of the system. Because of that, we defined T3R clones in such a way that they are directed towards finding clones that are worth this tradeoff.

@@

CloneRefactor uses JavaParser \cite{tomassetti2017javaparser}. JavaParser has a built-in symbol solver, which can automatically resolve types, method calls and variables. However, because of this, CloneRefactor does require all libraries that the software project requires (apart from the Java Standard Library). If these are not available, CloneRefactor will estimate types on basis of the information that is available.

@@

In this section, we first assess a set of modern clone detection tools for their applicability to this domain. Next, we introduce our own tool geared towards automatic clone refactoring, CloneRefactor.%Two surveys of modern clone detection tools \cite{sheneamer2016survey, svajlenko2014evaluating} together show an overview of the most-popular clone detection tools up until 2016. \todo{what is insufficient about these two? what triggered the need to reassess clone detectors? where are the four criteria coming from?}

\subsection{Survey on Clone Detection Tools}
\label{ch:tool-overview}
We conducted a short survey on (recent) clone detection tools that we could use to analyze refactoring possibilities. The results of our survey are displayed in table~\ref{table:dettools}. We chose a set of tools that are open source and can analyze a popular object-oriented programming language. Next, we formulate the following four criteria by which we analyze these tools:
\begin{enumerate}
    \item \textbf{Should find clones in any context.} Some tools only find clones in specific contexts, such as only method-level clones. We want to perform an analysis on all clones in projects to get a complete overview.
\item \textbf{Finds clone classes in control projects.} We assembled a number of control projects to assess the validity of clone detection tools. On basis of this, we checked whether clone detection tools can correctly find clones in diverse contexts.
\item \textbf{Can analyse resolved symbols.} When detecting clones for refactoring purposes, it is important that clone instances can be merged. Sometimes, textual equality between code fragments does not imply that these can be merged (this is described more elaborately in section \ref{sec:type1}). Because of this, we want to use a clone detection tool that can analyze such structures.
\item \textbf{Extensive detection configuration.} We aim to exclude expressions/statements from matching (more about our rationale in section~\ref{chap:clonetypes}). To achieve this, the tool needs to be able to allow those threshold changes. This can be either through simple changes of the source code, or by using some configuration file.
\end{enumerate}

\begin{table}[H]
 \begin{center}
  \caption{Our survey on clone detection tools.} \label{table:dettools}
  \medskip
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Clone Detection Tool} & \textbf{(1)} & \textbf{(2)} & \textbf{(3)} & \textbf{(4)} \\ \hline
Siamese \cite{ragkhitwetsagul2019siamese} &  &             &             & \checkmark            \\ \hline
NiCAD \cite{roy2008nicad, cordy2011nicad} & \checkmark                             & \checkmark            &             &             \\ \hline
CPD \cite{roy2009comparison} & \checkmark & \checkmark            &             &             \\ \hline
\begin{tabular}[c]{@{}l@{}}CCFinder \cite{kamiya2002ccfinder}\\ D-CCFinder \cite{livieri2007very}\end{tabular} & \checkmark  & \checkmark   &    &   \\ \hline
CCFinderSW \cite{semura2017ccfindersw}   & \checkmark     &             &             & \checkmark            \\ \hline
\begin{tabular}[c]{@{}l@{}}SourcererCC \cite{sajnani2016sourcerercc}\\ Oreo \cite{saini2018oreo}\end{tabular} & \checkmark    &             &             & \checkmark            \\ \hline
BigCloneEval \cite{svajlenko2016bigcloneeval}  & \checkmark  & \checkmark   &             &             \\ \hline
Deckard \cite{jiang2007deckard} & \checkmark   &             & \checkmark            &             \\ \hline
Scorpio \cite{higo2013revisiting, kamalpriya2017enhancing} & \checkmark   &     & \checkmark  & \checkmark   \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{CloneRefactor}
None of the state-of-the-art tools we identified implement all our criteria, so we decided to implement our own clone detection tool: CloneRefactor\footnote{CloneRefactor (WIP) is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}. This repository contains all scripts that were used to retrieve the data that is displayed in this paper.}. In this tool, we implemented both the literature clone types \cite{roy2007survey} and our refactoring-oriented clone types as described in section \ref{sec:rtypes}.

Our tool is based on the JavaParser library \cite{tomassetti2017javaparser}. This library can parse Java source code to an AST representation. This AST can then be analyzed, modified and eventually written back to source code. This allows us to perform AST-based clone detection and apply transformations to the AST based on the clones found.

CloneRefactor walks the AST and collects each declaration and statement (similar to the Scorpio clone detection tool \cite{higo2013revisiting}). It then builds a graph representation on basis of this AST, in which each declaration and statement becomes a node. This graph representation maps the following relations for each declaration and statement:

@@

\subsection{Thresholds}\label{chap:thresholds}
Thresholds are a tool that aid in the process of deciding whether a clone \textit{should} be refactored. Many clone detection tools focus on either the number of lines \cite{kamiya2002ccfinder, svajlenko2016bigcloneeval}, number of tokens \cite{roy2008nicad, sajnani2016sourcerercc, ragkhitwetsagul2019siamese} and/or number of nodes (declarations/statements) \cite{higo2013revisiting} to decide whether code fragments should be considered clones of each other. A comparison of clone detection tools by Bellon et al. \cite{bellon2007comparison} shows that most clone detection tools choose a minimum of \textbf{6} lines of code to be duplicate to consider code fragments a clone. The Scorpio clone detection tool uses this same number as minimum number of nodes \cite{higo2013revisiting}. Token based tools mostly operate on basis of a minimum of \textbf{50} tokens for fragments to be considered a clone \cite{sajnani2016sourcerercc}.

@@

We analyzed the context of clones in a large corpus of open source projects. For these experiments, we used our CloneRefactor tool. These experiments follow the structure of the context: The relation between clone instances is explained, measured and discussed in section \ref{chap:relationsinstances}; the location of clone instances is explained, measured and discussed in section \ref{chap:clonelocation}; the content of clone instances are explained, measured and discussed in section \ref{chap:clonecontents}.

\subsection{The corpus}\label{chap:corpus}
For our measurements we use a large corpus of open source projects \cite{githubCorpus2013}. This corpus has been assembled to contain relatively higher quality projects. Also, any duplicate projects were removed from this corpus. This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

As indicated in section \ref{chap:challenge} CloneRefactor requires all libraries of software projects we test. As these are not included in the used corpus \cite{githubCorpus2013}, we decided to filter the corpus to only include Maven projects. Maven is a build automation tool used primarily for Java, and works on basis of an \texttt{pom.xml} file to describe the projects' dependencies. As no \texttt{pom.xml} files are included in the corpus, we cloned the latest version of each project in the corpus. We then removed each project that has no \texttt{pom.xml} file. As a final step, we collected all dependencies for each project by using the \texttt{mvn dependency:copy-dependencies -DoutputDirectory=lib} Maven command, and removed each project for which not all dependencies were available (due to non-Maven dependencies being used or unsatisfiable dependencies being referenced in the \texttt{pom.xml} file).

@@

\begin{table}[H]
  \begin{center}
  \caption{General results for GitHub Java projects corpus \cite{githubCorpus2013}.} \label{table:general}
  \medskip
\begin{tabular}{|l|l|}
\hline
Amount of projects                                                                                      & 1,361      \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines (excluding\\whitespace, comments and newlines.)\end{tabular} & 1,414,996  \\ \hline
Amount of statements/declarations                                                                       & 1,212,189  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens (excluding\\whitespace, comments and newlines.)\end{tabular} & 11,643,194 \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{Clone detection results}
Currently, we have implemented two clone detection algorithms into CloneRefactor. The first one finds clones by comparing tokens (excluding whitespace, comments and newlines), equal to the definition of type 1 clones in literature \cite{roy2007survey}. The second algorithm implements our T1R, as explained in section \ref{sec:type1r}. The differences between the clones found for these algorithms is displayed in Table \ref{table:clonedet}.

@@

\subsection{Relations Between Clone Instances} \label{chap:relationsinstances}
When merging code clones in object-oriented languages, it is very important to consider the relation between clone instances. This relation has a big impact on how a clone should be refactored, in order to improve the software design in the process. In this section, we display measurements we conducted on the corpus introduced in section \ref{chap:corpus}. These measurements are based on an experiment by Fontana et al. \cite{fontana2015duplicated}, which we will briefly introduce in section \ref{chap:catcloneinstancerelations}. We use a vastly different setup, which is explained in section \ref{chap:oursetup}. We then show our results in section \ref{chap:ourmeasurements}.

\subsubsection{Categorizing Clone Instance Relations}\label{chap:catcloneinstancerelations}
Fontana et al. \cite{fontana2015duplicated} describe measurements on 50 open source projects on the relation of clone instances to each other. To do this, they first define several categories for the relation between clone instances in object-oriented languages. A few of these categories are shown in Figure \ref{fig:clonerelation}. These categories are as follows:
\begin{enumerate}
  \item \textbf{Same method}: All instances of the clone class are in the same method.
  \item \textbf{Same class}: All instances of the clone class are in the same class.
  \item \textbf{Superclass}: All instances of the clone class are children and parents of each other.
  \item \textbf{Ancestor class}: All instances of the clone class are superclasses except for the direct superclass.
  \item \textbf{Sibling class}: All instances of the clone class have the same parent class.
  \item \textbf{First cousin class}: All instances of the clone class have the same grandparent class.
\item \textbf{Common hierarchy class}: All instances of the clone class belong to the same hierarchy, but do not belong to any of the other categories.
\item \textbf{Same external superclass}: All instances of the clone class have the same superclass, but this superclass is not included in the project but part of a library.
\item \textbf{Unrelated class}: There is at least one instance in the clone class that is not in the same hierarchy.
\end{enumerate}

@@

We have looked into existing definitions for different types of clones \cite{roy2007survey} and proposed solutions for problems that these types have with regards to automated refactoring. We propose that fully qualified identifiers of method call signatures and type references should be considered instead of their plain text representation, to ensure refactorability. Furthermore, we propose that one should define thresholds for variability in variables, literals and method calls, in order to limit the number of parameters that the refactored unit shall have.

>>>>

\section{Introduction}
Duplication in source code is often seen as one of the most harmful types of technical debt. In Martin Fowler's ``Refactoring'' book~\cite{fowler1999refactoring}, he claims that \textit{``Number one in the stink parade is duplicated code. If you see the same code structure in more than one place, you can be sure that your program will be better if you find a way to unify them.''}.
% In this research, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve the maintainability of the refactored code.

@@

A survey by Roy et al.~\cite{roy2007survey} describes various ways in which clones can be identified. Most clone detection tools focus on finding clones that align with these definitions. In this paper, we outline challenges with these clone type definitions when considered in a refactoring context. We next propose solutions to these problems that would enable the detection of clones that should be refactored, rather than fragments of code that are just similar.

@@

Our end goal is to improve upon the current state-of-the-art in clone research~\cite{fontana2015duplicated, alwaqfi2017refactoring} by building a clone refactoring tool that automatically applies refactorings to a large percentage of clones found. The design decisions for this tool are made on basis of data gathered from a large corpus of software systems together with our own experience and findings from literature.

Sec.~\ref{chap:background} describes the background from literature that we used to conduct this study. Sec.~\ref{chap:clonedetection} presents our clone refactoring tool proposal. Sec.~\ref{chap:clonetypes} outlines challenges with clone type definitions and proposes solutions for them in a refactoring context. Sec.~\ref{chap:clonecontextexpl} outlines our analysis and measurements regarding the context of code clones. Sec.~\ref{chap:mergingdups} shows our measurements regarding the amount of clones that can be refactored through method extraction. Sec.~\ref{chap:threatstovalidity} shows the threats to validity of this study and Sec.~\ref{chap:conclusion} draws a conclusion based on our preliminary results.

@@

\subsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be considered for detection. We focus on clone classes, because of the advantages for refactoring. Clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics~\cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy~\cite{fontana2012duplicated}.

\subsection{Clone types}\label{chap:backgroundclonetypes}
In a 2007 survey by Roy et al.~\cite{roy2007survey} he defines several types of clones:\\
\textbf{Type 1:} Identical code fragments except for variations in whitespace (may also be variations in layout) and comments.\\
\textbf{Type 2:} Structurally/syntactically identical fragments except for variations in identifiers, literals, types, layout and comments.\\
\textbf{Type 3:} Copied fragments with further modifications. Statements can be changed, added or removed in addition to variations in identifiers, literals, types, layout, and comments.\\

A higher type of clone means that it is harder to detect. It also makes the clone harder to refactor, as more transformations would be required. Higher clone types also become more disputable whether they actually indicate a harmful anti-pattern (as not every clone is harmful~\cite{jarzabek2010clones, kapser2008cloning}).

@@

\subsection{Related work in clone refactoring tools}
The Duplicated Code Refactoring Advisor (DCRA) looks into refactoring opportunities for clone pairs~\cite{fontana2012duplicated, fontana2015duplicated}. DCRA only focuses on refactoring clone pairs, with the authors arguing that \textit{clone pairs are much easier to manage when considered singularly.} As intermediate steps, the authors measure a corpus of Java systems for some clone-related properties of the systems, like the relation (in terms of inheritance) between code fragments in a clone pair. We further look into these measurements in Sec.~\ref{chap:relationsinstances}.

Aries~\cite{higo2004aries, higo2008metric} focuses on the detection of refactorable clones. They do this based on the relation between clone instances through inheritance, similar to Fontana et al.~\cite{fontana2012duplicated}. Aries also proposes a refactoring: if two clone classes are siblings of each other (share the same superclass), they propose to perform ``Extract method'' and ``Pull up method'' sequentially. This tool only proposes the refactoring, and does not provide help in the process of applying the refactoring.

We investigated several research efforts that look into code clone refactoring~\cite{alwaqfi2017refactoring, chen2018clone, koni2001scenario}. However, all of these tools only support a subset of all harmful clones that are found. Also, these tools are limited to suggesting refactoring opportunities, rather than actually applying refactorings where suitable. Finally, all published approaches have limitations, such as false positives in their clone detection~\cite{chen2018clone} or being limited to clone pairs~\cite{higo2008metric}.

\section{Addressing problems with clone type definitions}\label{chap:clonetypes}
For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings (see Sec.~\ref{chap:backgroundclonetypes} for these definitions) to increase the chance that we can refactor the clone while improving the design.

\section{Shortcomings of clone types}
The clone definitions, as outlined in Sec.~\ref{chap:backgroundclonetypes}, allow reasoning about the duplication in a software system. Clones by these definitions can relatively easily and efficiently be detected. This has allowed for large scale analyses of duplication. However, these clone type definitions have shortcomings which make the clones detected in correspondence with these definitions less valuable for (automated) refactoring purposes.

@@

\subsection{Type 1 clones} \label{sec:type1}
Type 1 clones are \textit{identical clone fragments except for variations in whitespace and comments}~\cite{roy2007survey}. This allows for the detection of clones that are the result of copying and pasting existing code, along with other reasons why duplicates might get into a codebase.

@@

\subsection{Type 2 clones}\label{sec:type2}
Type 2 clones are \textit{structurally/syntactically identical fragments except for variations in identifiers, literals, types, layout and comments}~\cite{roy2007survey}. This definition allows for the reasoning about code fragments that were copied and pasted, and then slightly modified. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones. Merging such clones can be harmful instead of helpful.

@@

\subsection{Type 3 clones}\label{sec:type3}
Type 3 clones are \textit{copied fragments with further modification (having added, removed or changed statements)}~\cite{roy2007survey}. Detection of clones by this definition can be hard, as it may be hard to detect whether a fragment was copied in the first place if it was severely changed. Because of this, most clone detection implementations of type 3 clones work on basis of a similarity threshold~\cite{roy2008nicad,ragkhitwetsagul2019siamese,jiang2007deckard,semura2017ccfindersw}. This similarity threshold has been implemented in different ways: textual similarity (for instance using levenshtein distance)~\cite{lavoie2011automated}, token-level similarity~\cite{sajnani2016sourcerercc} of statement-level similarity~\cite{kamalpriya2017enhancing}.

@@

\subsubsection{T3R clones}\label{sec:type3r}
Type 3 clones allow any change in statements (added, removed and changed statements). Many clone detection tools have implemented this type by introducing a similarity threshold~\cite{roy2008nicad,ragkhitwetsagul2019siamese,jiang2007deckard,semura2017ccfindersw}. This similarity threshold entails that two fragments may differ by a certain percentage. This means that T3R allows the inclusion of a statement that is not detected by type 1 or 2 clone detection. When looking at how we can refactor a statement that is not included by one clone instance but is in another, we find that we require a conditional block to make up for the difference in statements. This is a tradeoff, as an added conditional block increases the complexity of the system. Because of that, we defined T3R clones in such a way that they are directed towards finding clones that are worth this tradeoff.

@@

CloneRefactor uses JavaParser~\cite{tomassetti2017javaparser}. JavaParser has a built-in symbol solver, which can automatically resolve types, method calls and variables. However, because of this, CloneRefactor does require all libraries that the software project requires (apart from the Java Standard Library). If these are not available, CloneRefactor will estimate types on basis of the information that is available.

@@

In this section, we first assess a set of modern clone detection tools for their applicability to this domain. Next, we introduce our own tool geared towards automatic clone refactoring, CloneRefactor.%Two surveys of modern clone detection tools~\cite{sheneamer2016survey, svajlenko2014evaluating} together show an overview of the most-popular clone detection tools up until 2016. \todo{what is insufficient about these two? what triggered the need to reassess clone detectors? where are the four criteria coming from?}

\subsection{Survey on Clone Detection Tools}
\label{ch:tool-overview}
We conducted a short survey on (recent) clone detection tools that we could use to analyze refactoring possibilities. The results of our survey are displayed in table~\ref{table:dettools}. We chose a set of tools that are open source and can analyze a popular object-oriented programming language. Next, we formulate the following four criteria by which we analyze these tools:
\begin{enumerate}
    \item \textbf{Should find clones in any context.} Some tools only find clones in specific contexts, such as only method-level clones. We want to perform an analysis on all clones in projects to get a complete overview.
\item \textbf{Finds clone classes in control projects.} We assembled a number of control projects to assess the validity of clone detection tools. On basis of this, we checked whether clone detection tools can correctly find clones in diverse contexts.
\item \textbf{Can analyse resolved symbols.} When detecting clones for refactoring purposes, it is important that clone instances can be merged. Sometimes, textual equality between code fragments does not imply that these can be merged (this is described more elaborately in Sec.~\ref{sec:type1}). Because of this, we want to use a clone detection tool that can analyze such structures.
\item \textbf{Extensive detection configuration.} We aim to exclude expressions/statements from matching (more about our rationale in section~\ref{chap:clonetypes}). To achieve this, the tool needs to be able to allow those threshold changes. This can be either through simple changes of the source code, or by using some configuration file.
\end{enumerate}

\begin{table}[H]
 \begin{center}
  \caption{Our survey on clone detection tools.} \label{table:dettools}
  \medskip
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Clone Detection Tool} & \textbf{(1)} & \textbf{(2)} & \textbf{(3)} & \textbf{(4)} \\ \hline
Siamese~\cite{ragkhitwetsagul2019siamese} &  &             &             & \checkmark            \\ \hline
NiCAD~\cite{roy2008nicad, cordy2011nicad} & \checkmark                             & \checkmark            &             &             \\ \hline
CPD~\cite{roy2009comparison} & \checkmark & \checkmark            &             &             \\ \hline
\begin{tabular}[c]{@{}l@{}}CCFinder~\cite{kamiya2002ccfinder}\\ D-CCFinder~\cite{livieri2007very}\end{tabular} & \checkmark  & \checkmark   &    &   \\ \hline
CCFinderSW~\cite{semura2017ccfindersw}   & \checkmark     &             &             & \checkmark            \\ \hline
\begin{tabular}[c]{@{}l@{}}SourcererCC~\cite{sajnani2016sourcerercc}\\ Oreo~\cite{saini2018oreo}\end{tabular} & \checkmark    &             &             & \checkmark            \\ \hline
BigCloneEval~\cite{svajlenko2016bigcloneeval}  & \checkmark  & \checkmark   &             &             \\ \hline
Deckard~\cite{jiang2007deckard} & \checkmark   &             & \checkmark            &             \\ \hline
Scorpio~\cite{higo2013revisiting, kamalpriya2017enhancing} & \checkmark   &     & \checkmark  & \checkmark   \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{CloneRefactor}
None of the state-of-the-art tools we identified implement all our criteria, so we decided to implement our own clone detection tool: CloneRefactor\footnote{CloneRefactor (WIP) is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}. This repository contains all scripts that were used to retrieve the data that is displayed in this paper.}. In this tool, we implemented both the literature clone types~\cite{roy2007survey} and our refactoring-oriented clone types as described in Sec.~\ref{sec:rtypes}.

Our tool is based on the JavaParser library~\cite{tomassetti2017javaparser}. This library can parse Java source code to an AST representation. This AST can then be analyzed, modified and eventually written back to source code. This allows us to perform AST-based clone detection and apply transformations to the AST based on the clones found.

CloneRefactor walks the AST and collects each declaration and statement (similar to the Scorpio clone detection tool~\cite{higo2013revisiting}). It then builds a graph representation on basis of this AST, in which each declaration and statement becomes a node. This graph representation maps the following relations for each declaration and statement:

@@

\subsection{Thresholds}\label{chap:thresholds}
Thresholds are a tool that aid in the process of deciding whether a clone \textit{should} be refactored. Many clone detection tools focus on either the number of lines~\cite{kamiya2002ccfinder, svajlenko2016bigcloneeval}, number of tokens~\cite{roy2008nicad, sajnani2016sourcerercc, ragkhitwetsagul2019siamese} and/or number of nodes (declarations/statements)~\cite{higo2013revisiting} to decide whether code fragments should be considered clones of each other. A comparison of clone detection tools by Bellon et al.~\cite{bellon2007comparison} shows that most clone detection tools choose a minimum of \textbf{6} lines of code to be duplicate to consider code fragments a clone. The Scorpio clone detection tool uses this same number as minimum number of nodes~\cite{higo2013revisiting}. Token based tools mostly operate on basis of a minimum of \textbf{50} tokens for fragments to be considered a clone~\cite{sajnani2016sourcerercc}.

@@

We analyzed the context of clones in a large corpus of open source projects. For these experiments, we used our CloneRefactor tool. These experiments follow the structure of the context: The relation between clone instances is explained, measured and discussed in Sec.~\ref{chap:relationsinstances}; the location of clone instances is explained, measured and discussed in Sec.~\ref{chap:clonelocation}; the content of clone instances are explained, measured and discussed in Sec.~\ref{chap:clonecontents}.

\subsection{The corpus}\label{chap:corpus}
For our measurements we use a large corpus of open source projects~\cite{githubCorpus2013}. This corpus has been assembled to contain relatively higher quality projects. Also, any duplicate projects were removed from this corpus. This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

As indicated in Sec.~\ref{chap:challenge} CloneRefactor requires all libraries of software projects we test. As these are not included in the used corpus~\cite{githubCorpus2013}, we decided to filter the corpus to only include Maven projects. Maven is a build automation tool used primarily for Java, and works on basis of an \texttt{pom.xml} file to describe the projects' dependencies. As no \texttt{pom.xml} files are included in the corpus, we cloned the latest version of each project in the corpus. We then removed each project that has no \texttt{pom.xml} file. As a final step, we collected all dependencies for each project by using the \texttt{mvn dependency:copy-dependencies -DoutputDirectory=lib} Maven command, and removed each project for which not all dependencies were available (due to non-Maven dependencies being used or unsatisfiable dependencies being referenced in the \texttt{pom.xml} file).

@@

\begin{table}[H]
  \begin{center}
  \caption{General results for GitHub Java projects corpus~\cite{githubCorpus2013}.} \label{table:general}
  \medskip
\begin{tabular}{|l|l|}
\hline
Amount of projects                                                                                      & 1,361      \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines (excluding\\whitespace, comments and newlines.)\end{tabular} & 1,414,996  \\ \hline
Amount of statements/declarations                                                                       & 1,212,189  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens (excluding\\whitespace, comments and newlines.)\end{tabular} & 11,643,194 \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{Clone detection results}
Currently, we have implemented two clone detection algorithms into CloneRefactor. The first one finds clones by comparing tokens (excluding whitespace, comments and newlines), equal to the definition of type 1 clones in literature~\cite{roy2007survey}. The second algorithm implements our T1R, as explained in Sec.~\ref{sec:type1r}. The differences between the clones found for these algorithms is displayed in Table \ref{table:clonedet}.

@@

\subsection{Relations Between Clone Instances} \label{chap:relationsinstances}
When merging code clones in object-oriented languages, it is very important to consider the relation between clone instances. This relation has a big impact on how a clone should be refactored, in order to improve the software design in the process. In this section, we display measurements we conducted on the corpus introduced in Sec.~\ref{chap:corpus}. These measurements are based on an experiment by Fontana et al.~\cite{fontana2015duplicated}, which we will briefly introduce in Sec.~\ref{chap:catcloneinstancerelations}. We use a vastly different setup, which is explained in Sec.~\ref{chap:oursetup}. We then show our results in Sec.~\ref{chap:ourmeasurements}.

\subsubsection{Categorizing Clone Instance Relations}\label{chap:catcloneinstancerelations}
Fontana et al.~\cite{fontana2015duplicated} describe measurements on 50 open source projects on the relation of clone instances to each other. To do this, they first define several categories for the relation between clone instances in object-oriented languages. A few of these categories are shown in Figure \ref{fig:clonerelation}. These categories are as follows:
\begin{enumerate}
  \item \textbf{Same method}: All instances of the clone class are in the same method.
  \item \textbf{Same class}: All instances of the clone class are in the same class.
  \item \textbf{Superclass}: All instances of the clone class are children and parents of each other.
  \item \textbf{Ancestor class}: All instances of the clone class are superclasses except for the direct superclass.
  \item \textbf{Sibling class}: All instances of the clone class have the same parent class.
  \item \textbf{First cousin class}: All instances of the clone class have the same grandparent class.
\item \textbf{Common hierarchy class}: All instances of the clone class belong to the same hierarchy, but do not belong to any of the other categories.
\item \textbf{Same external superclass}: All instances of the clone class have the same superclass, but this superclass is not included in the project but part of a library.
\item \textbf{Unrelated class}: There is at least one instance in the clone class that is not in the same hierarchy.
\end{enumerate}

@@

We have looked into existing definitions for different types of clones~\cite{roy2007survey} and proposed solutions for problems that these types have with regards to automated refactoring. We propose that fully qualified identifiers of method call signatures and type references should be considered instead of their plain text representation, to ensure refactorability. Furthermore, we propose that one should define thresholds for variability in variables, literals and method calls, in order to limit the number of parameters that the refactored unit shall have.
